A Generic and Efficient E-field Parallel Imaging Correlator for Next-Generation Radio Telescopes

### 2016-08-19 imaging_demo/ms.tex
Update abstract, Intro with Adam's suggestions
From 0c31757 (mnras):
====================

In ms.tex,

(*) Update abstract and Introduction with Adam's suggestions

% Abstract of the paper
\begin{abstract}
Modern radio telescopes are favouring densely packed array layouts with large numbers of antennas ($\Nant\gtrsim 1000$). Since the complexity of traditional correlators scales as $\mathcal{O}(\Nant^2)$, there will be a steep cost for realizing the full imaging potential of these powerful instruments. Through our generic and efficient E-field Parallel Imaging Correlator (EPIC), we present the first software demonstration of a generalized direct imaging algorithm, namely, the Modular Optimal Frequency Fourier (MOFF) imager. It takes advantage of the multiplication-convolution theorem of Fourier transforms. Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregular layouts and heterogeneous arrays of antennas. We demonstrate its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to severe mis-estimates. EPIC is highly modular and parallelizable, implemented in object oriented Python, and publicly available. We have verified the images produced to be equivalent to those produced using traditional techniques to within a precision set by gridding coarseness. We have also validated our implementation on data observed with the Long Wavelength Array (LWA). Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying EPIC. Inherent availability of calibrated time-domain images on digitizer writeout time-scales and vastly lower I/O bandwidth relative to visibility-based systems will make it a prime candidate for transient searches of Fast Radio Bursts (FRB) as well as planetary and exoplanetary phenomena. 
\end{abstract}

@@

Radio astronomy is entering an era in which interferometers of hundreds to thousands of individual antennas are needed to achieve desired survey speeds. Nowhere is this more apparent than at radio frequencies below 1.4 GHz. The study of the history of hydrogen gas throughout the universe's evolution is pushing technology development towards arrays of low-cost antennas with large fields of view and densely packed layouts. Similarly, the search for transient objects and regular monitoring of the time-dependent sky is driving instruments in the same direction with the added requirement of fast read-outs. A number of new telescopes are being or were developed around the world based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).

>>>>

% Abstract of the paper
\begin{abstract}
Modern radio telescopes are favouring densely packed array layouts with large numbers of antennas ($\Nant\gtrsim 1000$). Since the complexity of traditional correlators scales as $\mathcal{O}(\Nant^2)$, there will be a steep cost for realizing the full imaging potential of these powerful instruments. Through our generic and efficient E-field Parallel Imaging Correlator (EPIC), we present the first software demonstration of a generalized direct imaging algorithm, namely, the Modular Optimal Frequency Fourier (MOFF) imager. It takes advantage of the multiplication-convolution theorem of Fourier transforms. Not only does it bring down the cost for dense layouts to $\mathcal{O}(\Nant\log_2\Nant)$ but can also image from irregular layouts and heterogeneous arrays of antennas. EPIC is highly modular and parallelizable, implemented in object oriented Python, and publicly available. We have verified the images produced to be equivalent to those produced using traditional techniques to within a precision set by gridding coarseness. We have also validated our implementation on data observed with the Long Wavelength Array (LWA). We have demonstrated its versatility to image with heterogeneous arrays by showing that it robustly estimates the input sky model while wrong assumptions about array homogeneity lead to systematic mis-estimates. Antenna layouts with dense filling factors consisting of a large number of antennas such as LWA, the Square Kilometre Array, Hydrogen Epoch of Reionization Array, and Canadian Hydrogen Intensity Mapping Experiment will gain significant computational advantage by deploying EPIC. Inherent availability of calibrated time-domain images on digitizer writeout time-scales and vastly lower I/O bandwidth relative to visibility-based systems will make it a prime candidate for transient searches of Fast Radio Bursts (FRB) as well as planetary and exoplanetary phenomena. 
\end{abstract}

@@

Radio astronomy is entering an era in which interferometers of hundreds to thousands of individual antennas are needed to achieve desired survey speeds. Nowhere is this more apparent than at radio frequencies below 1.4 GHz. The study of the history of hydrogen gas throughout the universe's evolution is pushing technology development towards arrays of low-cost antennas with large fields of view and densely packed layouts. Similarly, the search for transient objects and regular monitoring of the time-dependent sky is driving instruments in the same direction with the added requirement of fast read-outs. A number of new telescopes are being or were developed around the world based on this new paradigm, including the Hydrogen Epoch of Reionization Array\footnote{http://reionization.org} (HERA; \citealt{deb16}), the Murchison Widefield Array (MWA; \citealt{tin13,bow13}), the Donald~C.~Backer Precision Array for Probing the Epoch of Reionization (PAPER; \citealt{par10}), the LOw Frequency ARray (LOFAR; \citealt{van13}), the Canadian Hydrogen Intensity Mapping Experiment (CHIME; \citealt{ban14}), the Long Wavelength Array (LWA; \citealt{ell13}), and the low frequency Square Kilometer Array (SKA1-Low; \citealt{mel13}).
